\paperTitle

### 2014-05-12 doc/icml2014/directed.tex
minor

%Let us now generalize the intuitions of the directed grid to general graphs.
The intuitions of the directed grid generalize readily to general graphs.
% PL: this prose is redundant with property
%Note that in each of the above cases,
  %for any clique $\sC$, each hidden variable $h_i \in \sC$ has its own
  %view which is conditionally independent given the other hidden variables in
  %the clique.
%Fundamentally, algorithm takes advantage of having a view for every
  %hidden variable in a clique to help identify its hidden state.
The key property required for our algorithm to work is as follows:
%We capture this general property as follows:
\begin{property}(Exclusive views)
  \label{prop:exclusive-views}
A hidden clique $\sC \in \sG$ is said to possess the \emph{exclusive views property} if for
  every hidden variable $h_i \in \sC$, (i)
  there exists some observed variable,
  $x_{v}$ which is conditionally independent of the rest of the clique
  given $h_i$, and (ii) the conditional moments $\mOpp{v}{i} \eqdef
  \Pr(x_{v} \mid h_i)$ can be estimated consistently.
We call $x_v$ an exclusive view of $h_i$ in $\sC$.
If every clique has exclusive views, then we say that $\sG$ does as well.
\end{property}
% PL: to save space
%This property also extends to any connected subgraph $\sG' \in \sG$.
%For the rest of this section, we will only consider subgraphs which are
  %cliques. 
%In \sectionref{undirected}, we will show how the generalization
  %to subgraphs allows us to efficiently estimate the pseudo-likelihood.

@@

% DONE: move here because not central to story
\paragraph{Remarks.} We note that \propertyref{bottleneck} can be relaxed if some cliques
  share parameters.
In the directed grid model, we can identify the conditional moments $O$ from just
  $x^a_1, x^b_1$ and $x^a_2$.
  Therefore, we do not require that $h_2, h_3$ or $h_4$
  be bottlenecks, and can omit the observations $x^b_2, x^b_3$ and $x^b_4$.

The method also extends directly to case in which the observed variables
  are real valued, while the hidden variables remain discrete. 
In this scenario, the tensor factorization method of
  \citet{anandkumar13tensor} recovers the expected conditional means,
  $\mu_{vi} \eqdef \E[x_v | h_i]$ for an observed variable $x_v$ and
  hidden variable $h_i$.
\algorithmref{learnclique} and \algorithmref{directed} still apply and allow
  us to recover clique marginals $Z_\sC$.
%  \todo{make this a footnote if running out of space}
% PL: don't understand this comment
%In this case, we only require that the hidden variables in distinct
  %cliques share parameters.

@@

In the HMM (\figureref{examples-hmm}), the parameters
are $O \eqdef \Pr(x_i|h_i)$  and that $T \eqdef \Pr(h_{i+1} | h_i)$
for all $i$. % (i.e. we have parameter sharing).
While the first and last hidden variables $h_1, h_M$ in the
  sequence are not bottlenecks, they still have exclusive views ($x_1$ and
  $x_M$ respectively) whose parameters we know because they share
  parameters, $O$.
%We first use the bottleneck $h_2$ with views $x_1, x_2, x_3$ to estimate $O$,
%and then recover $\pi$ from $\{h_1\}$ and recover $T$ from the clique $\{h_{1}, h_{2}\}$.

@@

In the latent tree model (\figureref{examples-tree}), the parameters
are $\Pr(h_i) = \pi$, $\Pr(h_i | h_1) = T$ for $i \in \{2,3,4\}$
and $\Pr(x^a_i | h_i) = \Pr(x^b_i | h_i) = O$ for $i \in \{2,3,4\}$.
Note that while $h_1$ is not directly connected to an observed variable,
  it is still a bottleneck, with views $x^a_2, x^a_3, x^a_4$.
We can recover $T$ from the clique $\{h_1, h_2\}$ by using views $x^a_2$
  (exclusive to $h_2$) and $x^a_3$ (exclusive to $h_1$).
  %Note that while
  %$x^a_3$ is also a view for $h_2$, $x^a_3$ is independent of $h_2$ given
  %$h_1$.

@@

The simplest example of a model that cannot be recovered by our algorithm is
  a mixture model with a single view, i.e. \figureref{three-view} with
  just the variables $h_1$ and $x_1$.
However, without further assumptions the model itself is not
  identifiable\footnote{To see this, observe that probability-mass
  can be arbitrarily exchanged between $\Pr(h_1)$ and $\Pr(x_1 | h_1)$ while preserving the marginals $\Pr(x_1)$}.
A class of binary-valued noisy-or networks
(see \figureref{examples-noisy-or} for an example)
which do not possess the
  bottleneck property are nonetheless identifiable;
  see \citet{halpern13noisyor} for an algorithm.

@@

To apply $\TensorFactorize$, we need that each $\mOpp{v}{i}$ has full
  column rank and that $\pi^{(i)} \eqdef \Pr(h_i) \succ 0$.
% But $\mOpp{v}{i}$ is a product of tensors on the path from $h_i$ to
%   $x_v$, marginalizing out the rest of the graph.
% To make this property more interpretable, let us assume the following
%   property which we prove to be sufficient in
%   \appendixref{assumption-proof}:
% \begin{assumption} 
% \label{asm:full-rank-plus}
% The marginal distribution of every latent variable has full support:
%   $\forall{h_i \in H},~ \mPi{i} \succ 0$.
% For every clique $\sC \in \sG$ (including ones with observed variables),
%   and every conditional distribution $\Pr(h_c \given \Pa(h_c))$, where
%   $h_c \in \sC$, every {\em tensor unfolding} has full column rank, $k$.
% \end{assumption}
% In our directed grid example (\sectionref{directedExample}), for the
%   clique $\sC = \{ h_1, h_2 \}$ this condition implies $\pi = \Pr(h_1)
%   \succ 0$ and $T = \Pr(h_2 \mid h_1)$ has column rank $k$; for the clique
%   $\sC = \{ h_1, x_1 \}$, it implies $O = \Pr(x_1 \mid h_1)$ has column
%   rank $k$.

>>>>

%Let us now generalize the intuitions of the directed grid to general graphs.
The intuitions of the directed grid generalize readily to general graphs.
% PL: this prose is redundant with property
%Note that in each of the above cases,
  %for any clique $\sC$, each hidden variable $h_i \in \sC$ has its own
  %view which is conditionally independent given the other hidden variables in
  %the clique.
%Fundamentally, algorithm takes advantage of having a view for every
  %hidden variable in a clique to help identify its hidden state.
The key property required for our algorithm to work is as follows:
%We capture this general property as follows:
\begin{property}[Exclusive views]
  \label{prop:exclusive-views}
A hidden clique $\sC \in \sG$ is said to possess the \emph{exclusive views property} if for
  every hidden variable $h_i \in \sC$, (i)
  there exists some observed variable,
  $x_{v}$ which is conditionally independent of the rest of the clique
  given $h_i$, and (ii) the conditional moments $\mOpp{v}{i} \eqdef
  \Pr(x_{v} \mid h_i)$ can be estimated consistently.
We call $x_v$ an exclusive view of $h_i$ in $\sC$.
If every clique has the exclusive views property, then we say that $\sG$ does as well.
\end{property}
% PL: to save space
%This property also extends to any connected subgraph $\sG' \in \sG$.
%For the rest of this section, we will only consider subgraphs which are
  %cliques. 
%In \sectionref{undirected}, we will show how the generalization
  %to subgraphs allows us to efficiently estimate the pseudo-likelihood.

@@

% DONE: move here because not central to story
\paragraph{Remarks.} We note that \propertyref{bottleneck} can be relaxed if some cliques
  share parameters.
For example, in the directed grid model, we can recover the conditional moments $O$ from
  $x^a_1, x^b_1$ and $x^a_2$ with $h_1$ as the bottleneck.
  Therefore, $h_2, h_3$ and $h_4$
  need not be bottlenecks, and we can omit the observations $x^b_2, x^b_3$ and $x^b_4$.

Our method also extends directly to case in which the observed variables
  are real-valued, while the hidden variables remain discrete. 
In this setting, the tensor factorization method of
  \citet{anandkumar13tensor} recovers the expected conditional means,
  $\mu_{vi} \eqdef \E(x_v | h_i) \in \Re^{d \times k}$ for each observed variable $x_v$ and
  hidden variable $h_i$.
\algorithmref{learnclique} and \algorithmref{directed} still apply and allow
  us to recover clique marginals $Z_\sC$.
%  \todo{make this a footnote if running out of space}
% PL: don't understand this comment
%In this case, we only require that the hidden variables in distinct
  %cliques share parameters.

@@

In the HMM (\figureref{examples-hmm}), the parameters
are $O \eqdef \Pr(x_i|h_i)$ and $T \eqdef \Pr(h_{i+1} | h_i)$
for all $i$. % (i.e. we have parameter sharing).
While the first and last hidden variables $h_1, h_M$ in the
  sequence are not bottlenecks, they still have exclusive views ($x_1$ and
  $x_M$, respectively)
  due to parameter sharing.
  %whose parameters we know because they share parameters, $O$.
%We first use the bottleneck $h_2$ with views $x_1, x_2, x_3$ to estimate $O$,
%and then recover $\pi$ from $\{h_1\}$ and recover $T$ from the clique $\{h_{1}, h_{2}\}$.

@@

In the latent tree model (\figureref{examples-tree}), the parameters
are $\pi \eqdef \Pr(h_i)$, $T \eqdef \Pr(h_i | h_1)$ for $i \in \{2,3,4\}$,
and $O \eqdef \Pr(x^a_i | h_i) = \Pr(x^b_i | h_i)$ for $i \in \{2,3,4\}$.
Note that while $h_1$ is not directly connected to an observed variable,
  it is still a bottleneck, with views $x^a_2, x^a_3, x^a_4$.
We can recover $T$ from the clique $\{h_1, h_2\}$ by using views $x^a_2$
  (exclusive to $h_2$) and $x^a_3$ (exclusive to $h_1$).
  %Note that while
  %$x^a_3$ is also a view for $h_2$, $x^a_3$ is independent of $h_2$ given
  %$h_1$.

@@

%In a certain sense, having three views is necessary for identifiability;
%two views is certainly not enough.
% PL: perhaps too simple, saving space
%The simplest example of a model that cannot be recovered by our algorithm is
%  a mixture model with a single view, i.e. \figureref{three-view} with
%  just the variables $h_1$ and $x_1$.
%However, without further assumptions the model itself is not
%  identifiable\footnote{To see this, observe that probability-mass
%  can be arbitrarily exchanged between $\Pr(h_1)$ and $\Pr(x_1 | h_1)$ while preserving the marginals $\Pr(x_1)$}.
%A class of binary-valued noisy-or networks
%(see \figureref{examples-noisy-or} for an example)
%which do not possess the
%  bottleneck property are nonetheless identifiable;
%  see \citet{halpern13noisyor} for an algorithm.
%Having three views is in some sense necessary for the uniqueness of tensor factorization.
There are certainly models which are identifiable but do not have exclusive views.
For example, \figureref{examples-noisy-or} shows
  a binary-valued noisy-or network which can be
  learned by the algorithm of \citet{halpern13noisyor},
  but does not satisfy the exclusive views property.

@@

To apply $\TensorFactorize$, we need that each $\mOpp{v}{i}$ has full
  column rank and that each $\pi^{(i)} \eqdef \Pr(h_i) \succ 0$.
% But $\mOpp{v}{i}$ is a product of tensors on the path from $h_i$ to
%   $x_v$, marginalizing out the rest of the graph.
% To make this property more interpretable, let us assume the following
%   property which we prove to be sufficient in
%   \appendixref{assumption-proof}:
% \begin{assumption} 
% \label{asm:full-rank-plus}
% The marginal distribution of every latent variable has full support:
%   $\forall{h_i \in H},~ \mPi{i} \succ 0$.
% For every clique $\sC \in \sG$ (including ones with observed variables),
%   and every conditional distribution $\Pr(h_c \given \Pa(h_c))$, where
%   $h_c \in \sC$, every {\em tensor unfolding} has full column rank, $k$.
% \end{assumption}
% In our directed grid example (\sectionref{directedExample}), for the
%   clique $\sC = \{ h_1, h_2 \}$ this condition implies $\pi = \Pr(h_1)
%   \succ 0$ and $T = \Pr(h_2 \mid h_1)$ has column rank $k$; for the clique
%   $\sC = \{ h_1, x_1 \}$, it implies $O = \Pr(x_1 \mid h_1)$ has column
%   rank $k$.
